---
title: "Project"
author: "Abhishek Anand, Mehul Jindal, Rosalia Giaccone"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/aanan/Documents/HUST Academics/Fall 2022/502/Project/") # change this to the folder that has your project files
library(Hmisc)
library(Rmisc)
library(psych)
library(moments)
library(VIM, quietly = T)
library(mice)
library(corrplot)
library(rio)
library(ggplot2)
library(cocor)
library(ppcor)
library(MOTE)
library(pwr)
library(papaja)
library(data.table)
library(MASS)
library(reshape2)
library(reshape)
library(tidyverse)
library(ez)
library(knitr)
library(dplyr)
library(statsr)
```


```{r, echo=FALSE, warning=FALSE, include=FALSE}
cleanup <- theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(), 
                panel.background = element_blank(), 
                axis.line.x = element_line(color = "black"),
                axis.line.y = element_line(color = "black"),
                legend.key = element_rect(fill = "white"),
                text = element_text(size = 10))
```


# Personal Vehicles Pricing and Features

## Introduction

```{r echo=FALSE}
Cars_Data <- readxl::read_excel("Cars_Data.xls") # Reading XLS excel file
Cars_Data <- rename(Cars_Data, "Sedan" = "Small/Sporty/ Compact/Large Sedan")
Cars_Data <- rename(Cars_Data, "Sports_Car" = "Sports Car")
Cars_Data <- rename(Cars_Data, "Retail_Price" = "Retail Price")
Cars_Data <- rename(Cars_Data, "Dealer_Cost" = "Dealer Cost")
Cars_Data <- rename(Cars_Data, "Engine_Size" = "Engine Size (l)")
Cars_Data <- rename(Cars_Data, "City_MPG" = "City MPG")
Cars_Data <- rename(Cars_Data, "Hwy_MPG" = "Hwy MPG")
Cars_Data <- rename(Cars_Data, "Wheel_Base" = "Wheel Base")
```

The Cars and Trucks Data lists specifications for 428 different vehicles from the year 2004, including factors related to the type of vehicle, prices, size, functionality, and fuel efficiency.

The data is accessible via the following link: http://jse.amstat.org/datasets/04cars.dat.txt

The variables contained in the data include the following:

- Type of Vehicle – the variants of vehicles named include the following: Sports, SUV, Wagon, Minivan, Pickup, AWD, RWD
- Dealer cost – the amount dealers pay manufacturers to buy the vehicle
- Engine size (liters)
- Number of Cylinders (=-1 if rotary engine)
- Horsepower
- City Miles per Gallon
- Highway Miles per Gallon
- Weight (pounds)
- Wheel Base (inches)
- Length (inches)
- Width (inches)


```{r echo=FALSE, results='hide'}
dim(Cars_Data)
```


```{r echo=FALSE, results='hide'}
colnames(Cars_Data)
```

### As Numeric

```{r}
cols.num <- c("City_MPG", "Hwy_MPG", "Weight", "Wheel_Base", "Len", "Width")
Cars_Data[cols.num] <- sapply(Cars_Data[cols.num], as.numeric)
```

## Missing Data

```{r Missing Data, echo=FALSE, results='hide'}
percentmiss <- function(x){sum(is.na(x))/length(x)*100}
missing <- apply(Cars_Data, 1, percentmiss)
table(missing)
```

```{r}
noerr <- Cars_Data
```

For data accuracy, we will replace data using built in software for anything at or below 10% missing data. We will ignore any entries missing more than 10%.

```{r}
replace_rows <- subset(noerr, missing <= 10)
noreplace_rows <- subset(noerr, missing >10)
```

Below we check for MNAR,

```{r}
apply(replace_rows,2,percentmiss)
```

```{r}
replace_columns <- replace_rows[, -c(1)]
noreplace_columns <- replace_rows[,c(1)]
```

```{r}
temp_no_miss <- mice(replace_columns, seed = 7)

nomiss <- complete(temp_no_miss, 1)

all_columns <- cbind(noreplace_columns,nomiss)
summary(all_columns)
```

No more missing values!

### Outliers

```{r}
mahal <- mahalanobis(all_columns[ , -1],
                    colMeans(all_columns[ , -1], na.rm=TRUE),
                    cov(all_columns[ , -1], use ="pairwise.complete.obs"),
                    tol=1e-30)
cutmahal <- qchisq(1-.001, ncol(all_columns[,-1]))

badmahal <- as.numeric(mahal >cutmahal)
table(badmahal)
```

```{r}
model1 <- lm(Retail_Price ~ Dealer_Cost + HP
                            + Engine_Size + Cyl
                            + City_MPG + Hwy_MPG 
                            + Weight + Wheel_Base
                            + Len + Width
                            + Sedan + Sports_Car
                            + SUV + Wagon
                            + Minivan + Pickup
                            + AWD + RWD,
          data = all_columns)
```

```{r}
k <- 18
leverage <- hatvalues(model1)
cutleverage <- (2*k+2)/nrow(all_columns)
badleverage <- as.numeric(leverage > cutleverage)
table(badleverage)
```

```{r}
cooks <- cooks.distance(model1)
cutcooks <- 4 / (nrow(all_columns) - k - 1)
badcooks <- as.numeric(cooks > cutcooks)
table(badcooks)
```

```{r}
totalout <- badmahal + badleverage + badcooks
table(totalout)
```

We see that there are 8 cars that are outliers based on all three methods, and 7 based on at least two methods. Hence, we will remove these.

```{r}
noout <- subset(all_columns,totalout<2)
onlyout <- subset(all_columns, totalout>=2)
```

We have removed outliers.

```{r}
model2 <- lm(Retail_Price ~ Dealer_Cost + HP
                            + Engine_Size + Cyl
                            + City_MPG + Hwy_MPG 
                            + Weight + Wheel_Base
                            + Len + Width
                            + Sedan + Sports_Car
                            + SUV + Wagon
                            + Minivan + Pickup
                            + AWD + RWD,
          data = all_columns)
```

## Assumptions

### Additivity

```{r}
summary(model2, correlation = TRUE)
```

There are some factors that are strongly correlated!

```{r}
corrplot(cor(noout[,-c(1,22)]))
```

Here, we see City_MPG and Hwy_MPG are highly correlated. We will only use Hwy_MPG in our modeling.

### Linearity

```{r linearity}
standardized <- rstudent(model2)
fitted <- scale(model2$fitted.values)
```

```{r}
{qqnorm(standardized)
abline(0,1)}
```

Linearity met.

### Normality

```{r normality}
hist(standardized)
```

Normality met.

### Homogeneity and Homoscedasticity

```{r homogs}
{plot(fitted, standardized)
abline(0,0)
abline(v = 0)}
```

## Generate Categorical Data

```{r}
noout$Weighted_MPG <- (noout$City_MPG + noout$Hwy_MPG)/2
summary(noout$Weighted_MPG)
noout$MPG_Cat <- as.factor(ifelse(noout$Weighted_MPG < 21, "Low_MPG",
                            ifelse(noout$Weighted_MPG > 25, "High_MPG","Medium_MPG")))
summary(noout$MPG_Cat)
```

```{r}
noout$mpg_diff <- noout$Hwy_MPG - noout$City_MPG

noout_v2 <- melt(noout, id.vars=c("Vehicle Name","AWD","RWD","Retail_Price",
                                  "Dealer_Cost","Engine_Size", "Cyl","HP","City_MPG",
                                  "Hwy_MPG","Weight","Wheel_Base","Len","Width","Weighted_MPG",
                                  "MPG_Cat","mpg_diff"))
noout_v2 <- subset(noout_v2, noout_v2$value==1)
```

## Summary Stats

```{r}
kable(summary(noout))
```

```{r}
kable(apply(noout,2,sd))
```


## Plots

```{r}
noout%>%
  ggplot(aes(Retail_Price))+
  geom_histogram(bins=30)+
  labs(title = 'Distribution of Retail Prices',
       y='Frequency',
       x='Retail Price')+
  cleanup +
  coord_cartesian(ylim = c(0,50))
```

```{r}
noout%>%
  ggplot(aes(Dealer_Cost, Retail_Price))+
  geom_point(alpha = 0.1) +
  #geom_histogram(bins=75)+
  labs(title = 'Retail Price and Dealer Cost',
       y='Retail Price',
       x='Dealer Cost')+
  cleanup #+
  #coord_cartesian(xlim = c(1,100), ylim = c(0,4000000))
#+ geom_smooth(method = 'lm', se=FALSE, color='navyblue')
```

```{r}
noout%>%
  ggplot(aes(Engine_Size, Retail_Price))+
  geom_point(alpha = 0.1) +
  #geom_histogram(bins=75)+
  labs(title = 'Retail Price by Engine Size',
       y='Retail Price',
       x='Engine Size')+
  cleanup +
  #coord_cartesian(xlim = c(1,100), ylim = c(0,4000000))
 geom_smooth(method = 'lm', se=FALSE, color='lightgreen')
```

```{r}
noout%>%
  ggplot(aes(Weighted_MPG, Retail_Price))+
  geom_point(alpha = 0.1) +
  #geom_histogram(bins=75)+
  labs(title = 'Retail Price by Mileage',
       y='Retail Price',
       x='Weighted MPG')+
  cleanup +
  #coord_cartesian(xlim = c(1,100), ylim = c(0,4000000))
 geom_smooth(method = 'lm', se=FALSE, color='lightgreen')
```

```{r}
noout%>%
  ggplot(aes(Weight, Retail_Price))+
  geom_point(alpha = 0.1) +
  #geom_histogram(bins=75)+
  labs(title = 'Retail Price by Vehicle Weight',
       y='Retail Price',
       x='Vehicle Weight')+
  cleanup +
  #coord_cartesian(xlim = c(1,100), ylim = c(0,4000000))
 geom_smooth(method = 'lm', se=FALSE, color='lightgreen')
```

```{r}
noout%>%
  ggplot(aes(HP, Retail_Price))+
  geom_point(alpha = 0.1) +
  #geom_histogram(bins=75)+
  labs(title = 'Retail Price by Vehicle Horsepower',
       y='Retail Price',
       x='Horsepower')+
  cleanup +
  #coord_cartesian(xlim = c(1,100), ylim = c(0,4000000))
 geom_smooth(method = 'lm', se=FALSE, color='lightgreen')
```

```{r}
noout%>%
  ggplot(aes(Retail_Price, reorder(MPG_Cat,
                                   FUN = median, Retail_Price,
                                   decreasing = TRUE)))+
  geom_boxplot()+
  labs(title = 'Difference in Prices by Mileage',
       x='Retail Price',
       y='Mileage Category')+
  cleanup #+
  #coord_cartesian(xlim = c(100000,4000000))
```

```{r}
noout_v2%>%
  ggplot(aes(Retail_Price, reorder(variable,
                                   FUN = median, Retail_Price,
                                   decreasing = TRUE)))+
  geom_boxplot()+
  labs(title = 'Difference in Prices by Vehicle Type',
       x='Retail Price',
       y='Vehicle Type')+
  cleanup #+
  #coord_cartesian(xlim = c(100000,4000000))
```

## noout_v3

```{r}
noout_v3 <- subset(noout_v2, select = -c(18:19))
noout_v3 <- melt(noout_v3, id.vars=c("Vehicle Name","Retail_Price",
                                  "Dealer_Cost","Engine_Size", "Cyl","HP","City_MPG",
                                  "Hwy_MPG","Weight","Wheel_Base","Len","Width","Weighted_MPG",
                                  "MPG_Cat","mpg_diff"))
noout_v3 <- subset(noout_v3, noout_v3$value==1)
```

```{r}
noout_v3%>%
  ggplot(aes(Retail_Price, reorder(variable,
                                   FUN = median, Retail_Price,
                                   decreasing = TRUE)))+
  geom_boxplot()+
  labs(title = 'Difference in Prices by Drive Power (Torque)',
       x='Retail Price',
       y='Drive Power')+
  cleanup #+
  #coord_cartesian(xlim = c(100000,4000000))
```


## Exploratory Data Analysis

## Hypothesis Test 1 - Engine_Size Difference in MPG Difference

```{r, warning=FALSE}
inference(y = Engine_Size, x = MPG_Cat, data = noout, statistic = "mean", type = "ht", null = 0, 
          alternative = "greater", method = "theoretical")
```

From output, we see that each of the three groups are unequal to each other. Hence, we reject null hypothesis.

From the box plots, we see that vehicles in High MPG category have the smallest Engine sizes, vehicles in the lowest MPG category have the largest engine sizes.

## Hypothesis Test 2 - Highway MPG vs City MPG

```{r}
inference(y = mpg_diff, data = noout, statistic ="mean", type="ci", method = "theoretical")
```

From the above output, we see that the 95% confidence interval does not contain 0. Hence, null hypothesis that the highway mpg and city mpg are equal is rejected. We see that, on average, highway mpg is 6.7 to 7.1 mpg higher than city mpg.

## Hypothesis Test 3

```{r}
inference(y = Retail_Price, x = variable, data = noout_v2, statistic = "mean", type = "ht", null = 0, 
          alternative = "greater", method = "theoretical")
```

We see that the prices for different vehicles are not the same. We reject null hypothesis of equal means in favor of the alternative.

## Modeling

```{r}
summary(model2)
```

## Because dealer cost and retail price are heavily collinear, we will remove dealer cost as an explanatory variable. It is almost as if we are trying to explain retail price variability using itself as a variable!

```{r}
model_best_v1 <- lm(Retail_Price ~ HP + Engine_Size + 
    Cyl + City_MPG + Hwy_MPG + Weight + Wheel_Base + Len + Width + 
    Sedan + Sports_Car + SUV + Wagon + Minivan + Pickup + AWD + 
    RWD, data = noout)
summary(model_best_v1)
```

## Hwy_MPG and City_MPG are highly collinear. Hence, we'll use just one - Hwy_MPG.

```{r}
model_best_v2 <- lm(Retail_Price ~ HP + Engine_Size + 
    Cyl + Hwy_MPG + Weight + Wheel_Base + Len + Width + 
    Sedan + Sports_Car + SUV + Wagon + Minivan + Pickup + AWD + 
    RWD, data = noout)
summary(model_best_v2)
```

## Engine Size, Cyl, and HP are highly collinear. We will only use HP.

```{r}
model_best_v3 <- lm(Retail_Price ~ HP +
                      Hwy_MPG + Weight + Wheel_Base + Len + Width + 
    Sedan + Sports_Car + SUV + Wagon + Minivan + Pickup + AWD + 
    RWD, data = noout)
summary(model_best_v3)
```

## Weight, Len, and Width are collinear. We will only Use Weight.

```{r}
model_best_v4 <- lm(Retail_Price ~ HP +
                      Hwy_MPG + Weight + Wheel_Base +
    Sedan + Sports_Car + SUV + Wagon + Minivan + Pickup + AWD + 
    RWD, data = noout)
summary(model_best_v4)
```
## Using p-value, we will do back-ward selection to get the best model:

```{r}
model_best_overall <- lm(Retail_Price ~ HP +
                      Hwy_MPG + Weight + Wheel_Base +
    Sedan + Sports_Car + SUV + Wagon + Minivan + Pickup + AWD + 
    RWD, data = noout)
summary(model_best_overall)
```


```{r}
summary(model_best_overall, correlation = TRUE)
```

## Assumptions

## Linearity

```{r linearity}
standardized <- rstudent(model_best_overall)
fitted <- scale(model_best_overall$fitted.values)
```

```{r}
{qqnorm(standardized)
abline(0,1)}
```

Linearity met.

### Normality

```{r normality}
hist(standardized)
```

Normality met.

### Homogeneity and Homoscedasticity

```{r homogs}
{plot(fitted, standardized)
abline(0,0)
abline(v = 0)}
```

## Final Model

$ RetailPrice = -12308 + 149*HP + 485*HwyMPG + 11*Weight-476*WheelBase+10074*Sedan+12325*SUV+8235*Wagon+6245*Minivan+3198*AWD+7254*RWD $

```{r}
HP = 0 # range from 93 - 450 in data
HwyMPG = 0 # range from 14 - 51 in data
Weight = 0 # range from 2035 - 6133 in data
WheelBase = 0 # range from 89 - 140 in data
Sedan = 0 # 0 or 1 #mutually exclusive with other car types
SUV = 0 # 0 or 1 #mutually exclusive with other car types
Wagon = 0 # 0 or 1 #mutually exclusive with other car types
Minivan = 0 # 0 or 1 #mutually exclusive with other car types
AWD = 0 # 0 or 1, mutually exclusive with RWD
RWD = 0 # 0 or 1, mutually exclusive with AWD

RetailPrice = -12308 + 149*HP + 485*HwyMPG + 11*Weight-476*WheelBase+10074*Sedan+12325*SUV+8235*Wagon+6245*Minivan+3198*AWD+7254*RWD 
```

## Prediction

Let us assume the first base case for a vehicle with the following:

A Sports Car with HP = 315, HwyMPG = 24, Weight = 3240, WheelBase = 93, AWD = 1, RWD = 0.

The price of this vehicle, according to our model, would be:

```{r}
HP = 315
HwyMPG = 24
Weight = 3240
WheelBase = 93
AWD = 1
RWD = 0

RetailPrice_Base = -12308 + 149*HP + 485*HwyMPG + 11*Weight-476*WheelBase+10074*Sedan+12325*SUV+8235*Wagon+6245*Minivan+3198*AWD+7254*RWD 
```

The base cost would be $40,837.

### Prediction 1

Compared to base vehicle, how would the price be different for a vehicle that weighed 1500 - i.e., severely under the weight range in data

```{r}
Weight = 1500

RetailPrice_LowWeight = -12308 + 149*HP + 485*HwyMPG + 11*Weight-476*WheelBase+10074*Sedan+12325*SUV+8235*Wagon+6245*Minivan+3198*AWD+7254*RWD
```

The cost of the low weight vehicle, given everything else remains the same as the base vehicle, decreases to $21,697 according to the model.

### Prediction 2

Compared to base vehicle, how would the price be different for a vehicle with a higher horsepower of 675, significantly above the current highest horsepower in data:

```{r}
Weight = 3240
HP = 675

RetailPrice_HighHP = -12308 + 149*HP + 485*HwyMPG + 11*Weight-476*WheelBase+10074*Sedan+12325*SUV+8235*Wagon+6245*Minivan+3198*AWD+7254*RWD
```

The cost of the high horsepower vehicle, given everything else remains the same as the base vehicle, increases to $94,477 according to this model!

### Prediction 3

Compared to base vehicle, how would the price be different for a vehicle with a significantly higher hwy mileage?

```{r}
HP = 315
HwyMPG = 75

RetailPrice_HighMPG = -12308 + 149*HP + 485*HwyMPG + 11*Weight-476*WheelBase+10074*Sedan+12325*SUV+8235*Wagon+6245*Minivan+3198*AWD+7254*RWD
```

The price of the high mileage vehicle, given everything else remains the same as the base vehicle, increases to $65,572 according to this model.

### Prediction 4

Compared to the base vehicle, how would the price be different for a vehicle with all three changes?

```{r}
Weight = 1500
HP = 675
HwyMPG = 75

RetailPrice_AllThree = -12308 + 149*HP + 485*HwyMPG + 11*Weight-476*WheelBase+10074*Sedan+12325*SUV+8235*Wagon+6245*Minivan+3198*AWD+7254*RWD
```

The price of a vehicle with low weight (1500), high HP (675) and high mileage (75) would be $100,072 according to this model!