---
title: "Project"
author: "Abhishek Anand"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/aanan/Documents/HUST Academics/Fall 2022/502/Project/") # change this to the folder that has your project files
library("Hmisc")
```

# Personal Vehicles Pricing and Features

## Introduction

```{r echo=FALSE}
Cars_Data <- readxl::read_excel("Cars_Data.xls") # Reading XLS excel file
```

The Cars and Trucks Data lists specifications for 428 different vehicles from the year 2004, including factors related to the type of vehicle, prices, size, functionality, and fuel efficiency.

The data is accessible via the following link: http://jse.amstat.org/datasets/04cars.dat.txt

The variables contained in the data include the following:

- Type of Vehicle – the variants of vehicles named include the following: Sports, SUV, Wagon, Minivan, Pickup, AWD, RWD
- Dealer cost – the amount dealers pay manufacturers to buy the vehicle
- Engine size (liters)
- Number of Cylinders (=-1 if rotary engine)
- Horsepower
- City Miles per Gallon
- Highway Miles per Gallon
- Weight (pounds)
- Wheel Base (inches)
- Length (inches)
- Width (inches)


```{r echo=FALSE, results='hide'}
dim(Cars_Data)
```


```{r echo=FALSE, results='hide'}
colnames(Cars_Data)
```

```{r Missing Data, echo=FALSE, results='hide'}
sum(is.na(Cars_Data)) # We see that there are no NAs.
```
## Exploratory Data Analysis

### Data Summary

```{r echo=FALSE, results='hide'}
describe(Cars_Data[10:20]) # Skip First 9 Columns that are just the type of vehicle [binary yes = 1, no = 0]
```

### Data Visualizations



### Inference


